Sweat electrolytes and CFTR activity in the era of CFTR modulators

### 2024-02-01 Edith Zemanick/Sweat Test/sweat_test_analysis.Rmd
Update sweat_test_analysis.Rmd

```{r}
# Visual inspection
ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red")
df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
kable(table(df$nacl_ratio_cat, df$Diagnosis))
```

@@

```{r}
set.seed(1017)
# K means
km_df <- df %>%
  select(Diagnosis, Na, Cl) %>%
  drop_na()
ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
  geom_point()
km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
km_df$Cluster <- factor(km$cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
centers <- data.frame(km$centers)
centers$Cluster <- rownames(km$centers)
centers$Cluster <- factor(centers$Cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
  geom_point(shape = 20, alpha = 0.2) +
  geom_point(data = centers, size = 5)
rownames(centers) <- NULL
centers <- centers %>% column_to_rownames("Cluster")
kable(centers, caption = "K-means centers", row.names = T)
kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
```

>>>>

```{r warning=FALSE}
# Visual inspection
ggplot(df, aes(x = Diagnosis, y = `Na/Cl Ratio`)) +
  geom_boxplot() +
  geom_hline(yintercept = 1, color = "red") +
  geom_hline(yintercept = 1.25, color = "red")
df$nacl_ratio_cat <- cut(df$`Na/Cl Ratio`, breaks = c(-Inf, 1, 1.25, Inf), right = F)
kable(table(df$nacl_ratio_cat, df$Diagnosis))
```

@@

### By Na and Cl

```{r warning=FALSE}
set.seed(1017)
# K means
km_df <- df %>%
  select(Diagnosis, Na, Cl) %>%
  drop_na()
ggplot(km_df, aes(x = Na, y = Cl, color = Diagnosis)) +
  geom_point()
km <- kmeans(km_df[, c("Na", "Cl")], centers = 3)
km_df$Cluster <- factor(km$cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
centers <- data.frame(km$centers)
centers$Cluster <- rownames(km$centers)
centers$Cluster <- factor(centers$Cluster,
  levels = 1:3,
  labels = c("High", "Medium", "Low")
)
ggplot(km_df, aes(x = Na, y = Cl, color = Cluster)) +
  geom_point(shape = 20, alpha = 0.2) +
  geom_point(data = centers, size = 5)
rownames(centers) <- NULL
centers <- centers %>% column_to_rownames("Cluster")
kable(centers, caption = "K-means centers", row.names = T)
kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
```

### By Na/Cl ratio

```{r warning=FALSE}
set.seed(1017)
# K means
km_df <- df %>%
  select(Diagnosis, `Na/Cl Ratio`) %>%
  drop_na()
km <- kmeans(km_df[, c("Na/Cl Ratio")], centers = 3)
km_df$Cluster <- factor(km$cluster,
  levels = 1:3,
  labels = c("Low","Medium","High")
)
centers <- data.frame(km$centers)
centers$Cluster <- rownames(km$centers)
centers$Cluster <- factor(centers$Cluster,
  levels = 1:3,
  labels = c("Low","Medium","High")
)
rownames(centers) <- NULL
centers <- centers %>% column_to_rownames("Cluster")
kable(centers, caption = "K-means centers", row.names = T)
kable(as.data.frame.matrix(table(km_df$Cluster, km_df$Diagnosis)), row.names = T)
```
